Performance Evaluation of Big Data Processing Strategies for Neuroimaging

### 2018-12-15 paper.tex
final final edits

In order to investigate how the different strategies impact processing, we 
selected a simple incrementation pipeline (Algorithm~\ref{alg:incrementation}) 
that consisted exclusively of map 
stages. A series of map-only stages would enable us to evaluate the effects of
in-memory computing when data locality is preserved. Incrementation was 
selected 
over other applications, such as binarization, as it ensured that a new image
was created at each step (i.e. no caching effects within the executing 
application). Each partitioned chunk was incremented by 1, in parallel, by
the applications. As incrementing the images is not a time consuming process,
we added a sleep delay to the incrementation stage to study the effects 
of tasks duration. The incremented chunks would be either 
maintained in-memory (Spark only) or saved to either tmpfs, local disk or 
Lustre (Spark and Nipype). Should more than a single iteration be requested, 
the incremented chunks would be incremented again and saved to the same file system. 
This would repeat until the number of requested iterations had elapsed. In all 
conditions, the first input chunks and final output chunks would reside on Lustre. 
We chose to perform our initial input/final output on Lustre as local storage is
typically only accessible to a user for the duration of the execution in HPC
environments.

@@

Fig.~\ref{fig:iterations} shows the difference between the different 
filesystem choices given the number of iterations. At 1 iteration, all 
filesystems behave the same, although the application was writing faster than the
disk bandwidth. This is because application data was not saturating the Page Cache
(transient phase). The page cache, on Zenith, occupies 40\% of 
total memory. With 192~GB of RAM on each node, 76.8~GB of 
dirty data could be held in a node's page cache at any given time.
As the total amount of data written by the 
application increases to 750~GB, there is a greater disparity between 
Lustre and in-memory (2.67~x slower, on average). Local disk 
performance, however, 
is still comparable to memory (1.38~x slower, on average). Despite local disk and 
Lustre both being in transient state, local disk encounters less contention 
than what would be found on Lustre.

@@

Spark currently doesn't include any active management of disk I/Os or 
page caches. We believe that it would be beneficial to extend it toward 
this direction, to increase the performance of operations where local 
storage has to be used, such as disk spills or shuffles. For instance, 
workflow-aware cache eviction policies that maximizes page cache usage 
for the workflows could be investigated.

>>>>

In order to investigate how the different strategies impact processing, we 
selected a simple incrementation pipeline (Algorithm~\ref{alg:incrementation}) 
that consisted exclusively of map 
stages. A series of map-only stages would enable us to evaluate the effects of
in-memory computing when data locality is preserved. Incrementation was 
selected 
over other applications, such as binarization, as it ensured that a new image
was created at each step (i.e. no caching effects within the executing 
application). Each partitioned chunk was incremented by 1, in parallel, by
the applications. As incrementing the images is not a time consuming process,
we added a sleep delay to the incrementation task to study the effects 
of tasks duration. The incremented chunks would be either 
maintained in-memory (Spark only) or saved to either tmpfs, local disk or 
Lustre (Spark and Nipype). Should more than a single iteration be requested, 
the incremented chunks would be incremented again and saved to the same file system. 
This would repeat until the number of requested iterations had elapsed. In all 
conditions, the first input chunks and final output chunks would reside on Lustre. 
We chose to perform our initial input/final output on Lustre as local storage is
typically only accessible to a user for the duration of the execution in HPC
environments.

@@

Fig.~\ref{fig:iterations} shows the difference between the different 
filesystem choices given the number of iterations. At 1 iteration, all 
filesystems behave the same, although the application was writing faster than the
disk bandwidth. This is because application data was not saturating the page cache
(transient phase). The page cache, on Zenith, occupies 40\% of 
total memory. With 192~GB of RAM on each node, 76.8~GB of 
dirty data could be held in a node's page cache at any given time.
As the total amount of data written by the 
application increases to 750~GB, there is a greater disparity between 
Lustre and in-memory (2.67~x slower, on average). Local disk 
performance, however, 
is still comparable to memory (1.38~x slower, on average). Despite local disk and 
Lustre both being in transient state, local disk encounters less contention 
than what would be found on Lustre.

@@

Spark does not currently include any active management of disk I/Os or 
page caches. We believe that it would be beneficial to extend it toward 
this direction, to increase the performance of operations where local 
storage has to be used, such as disk spills or shuffles. For instance, 
workflow-aware cache eviction policies that maximizes page cache usage 
for the workflows could be investigated.
